---
title: "Tutorial 3 - Markov chain Monte Carlo Calibration (Supplementary)"
author: "Nan-Hung Hsieh"
date: "2019/05/23 (update: `r Sys.Date()`)"
output: 
  html_document:
  fig_caption: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('..')
wd <- getwd()
knitr::opts_knit$set(root.dir =  wd)
```


```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(rstan)
library(bayesplot)
library(corrplot)
library(sensitivity)
library(pksensi)
theme_set(theme_light())
```

```{r, include=FALSE}
source("MCSim/function.R")
```

# 3 Tetrachloroethylene PBPK Model  

Prepare [**model**](https://raw.githubusercontent.com/nanhung/MCSim_under_R/master/modeling/perc/perc.model.R) (`perc.model.R`) and [**input**](https://raw.githubusercontent.com/nanhung/MCSim_under_R/master/modeling/perc/perc_mcmc.in.R) (`perc_mcmc.in.R`) files. This input file  contains MCMC simulation descriptions based on Monster et al. (1989) paper. The  study measured the TCE in blood (`C_ven`) and exhaled air (`C_exh_ug`) for a group of 6 subjects exposed to two different concentrations (72 and 144 ppm) of TCE in air.

## 3.1 Single level modeling

### 3.1.1 Single chain testing

**Modeling**

```{r}
model <- "perc.model.R"
input <- "perc_mcmc.in.R"   
set.seed(1234)
system.time(out_1 <- mcsim(model, input, dir = "modeling/perc"))
head(out_1)
```

**Diagnosis**

```{r}
str <- ceiling(nrow(out_1)/2) + 1
end <- nrow(out_1)
j <- c(str:end) # discard burn-in
plot(out_1$iter[j], out_1$sc_Vmax.1.[j], type = "l")
plot(out_1$iter[j], out_1$Km.1.[j], type = "l")
```

```{r}
out_1$sc_Vmax.1.[j] %>% density %>% plot()
out_1$Km.1.[j] %>% density() %>% plot()
```


```{r}
par(mfrow = c(1,2), mar = c(4,4,3,1))
acf(out_1$sc_Vmax.1.[j])
acf(out_1$Km.1.[j])
```

```{r}
cor <- cor(out_1$sc_Vmax.1.[j], out_1$Km.1.[j])
plot(out_1$sc_Vmax.1.[j], out_1$Km.1.[j],
     main = paste("r = ", round(cor, digits = 3)))
```

**Checking model fit**

1. Global evaluation

```{r}
chk_1 <- read.delim("MCMC.check.out")
tail(chk_1)
```

For multi level simulation, the "Level" column include **"population-experiments"**

```{r}
R_square <- cor(chk_1$Data, chk_1$Prediction)^2
plot(chk_1$Data, chk_1$Prediction, log = "xy",
     xlim = c(0.001, 1000), ylim = c(0.001, 1000),
     xlab = "Observation", ylab = "Prediction",
     main = paste("R2 =", round(R_square, digits = 3)))
abline(0,1)
```

2. Individual evaluation 

(1) Concentration in venous blood

```{r}
chk_1 %>% filter(Output_Var == "C_ven") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "Concentration (mg/L)", title = "Concentration in venous blood")
```

(2) Concentration in exhaled air

```{r}
chk_1 %>% filter(Output_Var == "C_exh_ug") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_wrap(~Simulation) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "Concentration (ug/L)", title = "Concentration exhaled air")
```


### 3.1.2 Checking convergence and model fit with multi chains

**Checking convergence**

```{r}
model <- "perc.model.R"
input <- "perc_mcmc.in.R"   
set.seed(2345)
out_1_2 <- mcsim(model, input, dir = "modeling/perc")
set.seed(3344)
out_1_3 <- mcsim(model, input, dir = "modeling/perc")
set.seed(4455)
out_1_4 <- mcsim(model, input, dir = "modeling/perc")
```

```{r}
x <- mcmc_array(list(out_1, out_1_2, out_1_3, out_1_4))
```

```{r}
parms <- c("sc_Vmax.1.", "Km.1.")
mcmc_dens_overlay(x[j,,], pars = parms)
mcmc_trace(x[j,,], pars = parms)
```

```{r}
monitor(x, digits = 4)
```

**Evaluation of model fit**

1. Setpoints analysis

```{r}
X <- x[j,,] %>% matrix(nrow = 16000) 
write.table(X, file = "setpts.out", row.names = F, sep = "\t")
X_setpts <- mcsim("perc.model.R", "perc_setpts.in.R", dir = "modeling/perc")
```

2. Tidy data with median and 95% confidence interval

```{r}
vars <- names(X_setpts)
index <- which(vars == "C_exh_ug_1.1" | vars == "C_ven_12.7")
X <- apply(X_setpts[index[1]:index[2]], 2, quantile,  c(0.5, 0.025, 0.975)) %>% t()
colnames(X) <- c("median", "LCL", "UCL")
df <- as.data.frame(X)
chk_1$med <- df$median
chk_1$LCL <- df$LCL
chk_1$UCL <- df$UCL
```

3. Plot result

```{r, warning=F}
chk_1 %>% filter(Output_Var == "C_ven") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_ribbon(aes(ymin = LCL, ymax = UCL), fill = "grey70", alpha = 0.5) +
  geom_line(aes(y = med)) +
  facet_wrap(~Simulation) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "Concentration (mg/L)", title = "Concentration in venous blood")
```

**Note:** The 95% confidence interval is too narrow for visual observing.

## 3.2 Hierarchical modeling

### 3.2.1 Single chain testing

**Modeling**

```{r}
model <- "perc.model.R"
input <- "perc_mcmc-hier.in.R"   
set.seed(1234)
system.time(out_2 <- mcsim(model, input, dir = "modeling/perc"))
head(out_2)
```

**Diagnosis**

(1) Trace plot

```{r}
str <- ceiling(nrow(out_2)/2) + 1
end <- nrow(out_2)
j <- c(str:end) # discard burn-in
plot(out_2$iter[j], out_2$sc_Vmax.1.[j], type = "l")
plot(out_2$iter[j], out_2$Km.1.[j], type = "l")
```

(2) Kernel plot

```{r}
out_2$sc_Vmax.1.[j] %>% density %>% plot()
out_2$Km.1.[j] %>% density() %>% plot()
```

(3) Auto-correlation plot

```{r}
par(mfrow = c(1,2), mar = c(4,4,3,1))
acf(out_2$sc_Vmax.1.[j])
acf(out_2$Km.1.[j])
```

(4) Pair plot

```{r}
cor_2 <- cor(out_2$Km.1.[j], out_2$sc_Vmax.1.[j])
plot(out_2$Km.1.[j], out_2$sc_Vmax.1.[j], 
     xlab = "Km", ylab = "Vmax",
     main = paste("r =", round(cor_2, digits = 3)))
```

**Evaluation of model fit**

```{r}
chk_2 <- read.delim("MCMC.check.out")
```

Compare the different between single and multi level

For multi level simulation, the "Level" column include **"population-individuals-experiments"**

```{r}
tail(chk_1)
tail(chk_2)
```


1. Global evaluation

```{r}
R_square <- cor(chk_2$Data, chk_2$Prediction)^2
plot(chk_2$Data, chk_2$Prediction, log = "xy",
     xlim = c(0.001, 1000), ylim = c(0.001, 1000),
     xlab = "Observation", ylab = "Prediction",
     main = paste("R2 =", round(R_square, digits = 3)))
abline(0,1)
```

2. Individual evaluation

```{r}
chk_2 %>% 
  separate(col = Level, c("Pop", "Subj", "Exp")) %>% 
  filter(Output_Var == "C_ven") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_line(aes(y = Prediction))+
  facet_grid(Exp~Subj) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "Concentration (mg/L)", title = "Concentration venous blood")
```


### 3.2.2 Checking convergence and model fit with multi chains

```{r}
model <- "perc.model.R"
input <- "perc_mcmc-hier.in.R"
set.seed(2345)
out_2_2 <- mcsim(model, input, dir = "modeling/perc")
set.seed(3344)
out_2_3 <- mcsim(model, input, dir = "modeling/perc")
set.seed(4455)
out_2_4 <- mcsim(model, input, dir = "modeling/perc")
```

```{r}
x <- mcmc_array(list(out_2, out_2_2, out_2_3, out_2_4))
```

**Diagnosis**

```{r, eval=F}
parms <- c("sc_Vmax.1.", "Km.1.", "sc_Vmax.1.1.", "Km.1.1.", "sc_Vmax.1.2.",
           "Km.1.2.", "sc_Vmax.1.3.", "Km.1.3.", "sc_Vmax.1.4.", "Km.1.4.", "sc_Vmax.1.5.",
           "Km.1.5.", "sc_Vmax.1.6.", "Km.1.6.")
mcmc_dens_overlay(x[j,,], pars = parms)
mcmc_trace(x[j,,], pars = parms)
```

```{r}
monitor(x, digits = 4)
```

**Evaluation of model fit**

1. Setpoints analysis

```{r}

X <- x[j,,] %>% matrix(nrow = 16000) 
write.table(X, file = "setpts.out", row.names = F, sep = "\t")
X_setpts <- mcsim("perc.model.R", "perc_setpts.in.R", dir = "modeling/perc")
```

2. Tidy data with median and 95% confidence interval

```{r}
vars <- names(X_setpts)
index <- which(vars == "C_exh_ug_1.1" | vars == "C_ven_12.7")
X <- apply(X_setpts[index[1]:index[2]], 2, quantile,  c(0.5, 0.025, 0.975)) %>% t()
colnames(X) <- c("median", "LCL", "UCL")
df <- as.data.frame(X)
chk_2$med <- df$median
chk_2$LCL <- df$LCL
chk_2$UCL <- df$UCL
```

3. Plot result

```{r, warning=F}
chk_2 %>% filter(Output_Var == "C_ven") %>%
ggplot(aes(x = Time, y = Data)) +
  geom_point() + 
  geom_ribbon(aes(ymin = LCL, ymax = UCL), fill = "grey70", alpha = 0.5) +
  geom_line(aes(y = med)) +
  facet_wrap(~Simulation) +
  scale_y_log10() + 
  labs(x = "Time (min)", y = "Concentration (mg/L)", title = "Concentration in venous blood")
```


## 3.3 Parallel computing

**How how many threads your CPU has?**

```{r, eval=F}
parallel::detectCores()
```


**Parallel computing**

Prepare the job file first, the example job file is put into "examples/perc_job.R".

Set `importEnv = T` to assign the random seed to the job. The `exportEnv` is used to assign the name of object to store the output data.


```{r, eval=F}
library(rstudioapi)
set.seed(1234)
jobRunScript(workingDir = getwd(), path = "examples/perc_job.R", importEnv = T, exportEnv = "job_1")
set.seed(2345)
jobRunScript(workingDir = getwd(), path = "examples/perc_job.R", importEnv = T, exportEnv = "job_2")
set.seed(3344)
jobRunScript(workingDir = getwd(), path = "examples/perc_job.R", importEnv = T, exportEnv = "job_3")
set.seed(4444)
jobRunScript(workingDir = getwd(), path = "examples/perc_job.R", importEnv = T, exportEnv = "job_4")
```

```{r, eval=F}
x <- mcmc_array(list(job_1$out_2, job_2$out_2, job_3$out_2, job_4$out_2))
```

```{r, include=FALSE}
clear()
```
