---
title: "Tutorial 2: Uncertainty and Sensitivity Analysis (Supplementary)"
author: "Nan-Hung Hsieh"
date: "2019/05/16 (update: `r Sys.Date()`)"
output: 
  html_document:
  fig_caption: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('..')
wd <- getwd()
knitr::opts_knit$set(root.dir =  wd)
```

# 4 Example 2: PERC-PBPK model 

```{r, include=FALSE}
source("MCSim/function.R")
```

The list of R packages should be installed first to do the following testing. 

```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(sensitivity)
library(pksensi)
```

## 4.1  Monte Carlo simulation in PERC-PBPK model

0. Prepare the model ("perc.model.R") and input ("perc_mtc.in.R") files. There are two simulations, which are inhalation of 72 and 144 ppm with 4 hour exposure.

1. Modeling

```{r, eval=T}
out <- mcsim(model = "perc.model.R", input = "perc_mtc.in.R", dir = "modeling/perc")
```

2. Data manipulate

```{r, eval=T}
vars <- names(out)
sim1.1 <- which(vars == "C_exh_ug_1.1" | vars == "C_exh_ug_1.10")
sim1.2 <- which(vars == "C_ven_1.1" | vars == "C_ven_1.8") 
sim2.1 <- which(vars == "C_exh_ug_2.1" | vars == "C_exh_ug_2.10")
sim2.2 <- which(vars == "C_ven_2.1" | vars == "C_ven_2.8")
```

3. Visualize

```{r, eval=T}
last.param <- which(names(out) == "Flow_pul")
par(mfrow = c(4, 4), mar = c(2,2,4,1))
for (i in 2:last.param){
  hist(out[,i], main = names(out[i]), xlab = "")
}
```

```{r, eval=T}
ggPK <- function(data, index, time, ...){
  X <- apply(data[,index[1]:index[2]], 2, quantile,  c(0.5, 0.025, 0.975))
  dat <- t(X)
  colnames(dat) <- c("median", "LCL", "UCL")
  df <- as.data.frame(dat)
  df$time <- time
  ggplot(df, aes(x = time, y = median)) +
    geom_ribbon(aes(ymin = LCL, ymax = UCL), fill = "grey70", alpha = 0.5) + 
    geom_line() +
    labs(...) +
    scale_y_log10()
}
```


```{r, eval=T}
t_exh_ug <- c(239.9, 245, 270, 360, 1320, 2760, 4260, 5700, 8580, 10020)
t_Ven <- c(239.9, 360, 1320, 2760, 4260, 5700, 8580, 10020)
ggPK(data = out, index = sim1.1, time = t_exh_ug, x = "Time (min)", y = "mg/l in the exhaled air")
ggPK(data = out, index = sim1.2, time = t_Ven, x = "Time (min)", y = "mg/l in the venous blood")
ggPK(data = out, index = sim2.1, time = t_exh_ug, x = "Time (min)", y = "mg/l in the exhaled air")
ggPK(data = out, index = sim2.2, time = t_Ven, x = "Time (min)", y = "mg/l in the venous blood")
```

## 4.2 Application of **pksensi** in Monte Carlo simulation 

1. Set parameter distribution (assign `parms`, `dist`, `q.qarg`)

```{r, eval=T}
parameters <- c("LeanBodyWt","Pct_M_fat", "Pct_LM_liv", "Pct_LM_wp",
                "Pct_Flow_fat", "Pct_Flow_liv", "Pct_Flow_pp",
                "PC_fat", "PC_liv", "PC_wp", "PC_pp", "PC_art",
                "Vent_Perf", "sc_Vmax", "Km", "Flow_pul")  
dist <- rep("Uniform", 16) # MCSim definition
q.arg<-list(list(50, 70),
            list(0.2, 0.3),
            list(0.03, 0.04),
            list(0.25, 0.3),
            list(0.06, 0.08),
            list(0.2, 0.3),
            list(0.2, 0.25),
            list(110, 150),
            list(5.0, 8.0),
            list(5.0, 8.5),
            list(1.6, 1.8),
            list(12, 15),
            list(0.8, 1.3),
            list(0.04, 0.06),
            list(7, 13),
            list(7.4, 7.6))
```

2. Set experiment time-points, output variables, and conditions 

```{r, eval=T}
times <- c(239.9, 245, 270, 360, 1320, 2760, 4260, 5700, 8580, 10020)
outputs <- c("C_exh_ug", "C_ven")
conditions <- c("InhMag = 72", "Period = 1e10", "Exposure = 240")
```                
                
3. Modeling

```{r, eval=T}
set.seed(2222)
y<-solve_mcsim(mName = "perc.model.R", params = parameters, vars = outputs,
               monte_carlo = 1000, dist = dist, q.arg = q.arg, time = times, 
               condition = conditions)
```

4. Visualization

```{r, fig.width=10, fig.height=3, eval=T}
par(mfrow = c(1, 2), mar = c(2,3,1,1))
pksim(y, vars = "C_exh_ug", log = T)
pksim(y, vars = "C_ven", log = T)
```

## 4.3 Application of pksensi in sensitivity analysis

1. Set parameter distribution (assign `parms`, `dist`, `q.qarg`)

```{r, eval=T}
parameters <- c("LeanBodyWt",
                "Pct_M_fat", "Pct_LM_liv", "Pct_LM_wp","Pct_Flow_fat", "Pct_Flow_liv", "Pct_Flow_pp",
                "PC_fat", "PC_liv", "PC_wp", "PC_pp", "PC_art",
                "Vent_Perf", 
                "sc_Vmax", "Km", 
                "Flow_pul")  
q <- "qunif"
q.arg <- list(list(50, 70),
            list(0.2, 0.3),
            list(0.03, 0.04),
            list(0.25, 0.3),
            list(0.06, 0.08),
            list(0.2, 0.3),
            list(0.2, 0.25),
            list(110, 150),
            list(5.0, 8.0),
            list(5.0, 8.5),
            list(1.6, 1.8),
            list(12, 15),
            list(0.8, 1.3),
            list(0.04, 0.06),
            list(7, 13),
            list(7.4, 7.6))
```

2. Generate parameter space

```{r, eval=T}
x <- rfast99(params = parameters, n = 2000, q = q, q.arg = q.arg, replicate = 1)
dim(x$a)
```


3. Set experiment time-points, output variables, and conditions 

```{r, eval=T}
times <- c(239.9, 245, 270, 360, 1320, 2760, 4260, 5700, 8580, 10020)
outputs <- c("C_exh_ug", "C_ven")
conditions <- c("InhMag = 72", "Period = 1e10", "Exposure = 240")
```

4. Modeling

```{r, eval=T}
y <- solve_mcsim(x, mName = "perc.model.R", params = parameters, vars = outputs,
                 time = times, condition = conditions)
```

5. Visualization & Decision


```{r, eval=T}
par(mfrow = c(4, 4), mar = c(2,2,4,1))
for(i in 1:16){
  hist(x$a[,,i])  
}
```


```{r fig.height=8, fig.width=13, eval=T}
plot(y)
```

```{r fig.height=8, fig.width=13, eval=T}
plot(y, vars = 2)
```


```{r fig.height=7, fig.width=13, eval=T}
heat_check(y) 
```

```{r, eval=T}
check(y)
```

# 5 Exercise

In the exercise 1 to 4, we'll use the `EB.model.R` to conduct Monte Carlo simulation and sensitivity analysis. 

## 5.1 Monte Carlo simulation

**Create the MCSim's input file ("EB_mtc.in.R") and run Monte Carlo simulation to conduct uncertainty analysis of EB concentration in blood (0 - 6 hr), the exposure condition is 100 ppm EB exposure for 4 hours.**

```{r, eval=T}
out <- mcsim(model = "EB.model.R", input = "EB_mtc.in.R", dir = "modeling/EB")
head(out)
```

Data manipulate

```{r, eval=T}
vars <- names(out)
index <- which(vars == "Cvtot_1.2" | vars == "Cvtot_1.61")
time <- seq(0.1, 6, 0.1)
MW <- 106.16 # g/mol
X <- apply(out[,index[1]:index[2]], 2, quantile,  c(0.5, 0.01, 0.99))
dat <- t(X) * MW * 1000
colnames(dat) <- c("median", "LCL", "UCL")
df <- as.data.frame(dat)
df$time <- time
head(df)
```

Plot

```{r, eval=T}
ggplot(df, aes(x = time, y = median)) +
    geom_ribbon(aes(ymin = LCL, ymax = UCL), fill = "grey70", alpha = 0.5) + 
    geom_line() 
```

## 5.2 Morris screening

**At the same exposure condition, conduct Morris elementary effects screening method to find the influential parameters for blood concentration during the 2-hr post exposure.**

1. Parameter setting

```{r, eval=T}
BW <- 0.043  

Pb <- 42.7
Pl <- 1.96
Pf <- 36.4
Pm <- 0.609
Pvrg <- 1.96
Ppu <- 1.96
Pbr <- 1.96

VmaxC <- 6.39
VmaxClu <- 13.4
VmaxCvr <- 17.4

factors <- c("BW", "Pb", "Pl", "Pf", "Pm", "Pvrg", "Ppu", "Pbr", "VmaxC", "VmaxClu", "VmaxCvr")

baseline <- c(BW, Pb, Pl, Pf, Pm, Pvrg, Ppu, Pbr, VmaxC, VmaxClu, VmaxCvr)
limit <- c(1.2, rep(2, 7), rep(4, 3))
binf <- baseline/limit
bsup <- baseline*limit

binf
bsup
```

2. Generate parameter matrix

```{r }
set.seed(1234)
x <- morris(model = NULL, factors = factors, r = 512,
            design = list(type = "oat", levels = 5, grid.jump = 3),
            binf = binf, bsup = bsup)
X <- cbind(1, x$X)
write.table(X, file = "setpts.out", row.names = F, sep = "\t")
head(X)
```

3. Modeling

```{r, eval=T}
out <- mcsim(model = "EB.model.R",  input = "EB_setpts.in.R", dir = "modeling/EB")
head(out)
```

4. Visualization

```{r, eval=T}
str <- which(names(out) == "Cvtot_1.1")
end <- which(names(out) == "Cvtot_1.5")
par(mfrow = c(2,3), mar = c(2,2,2,1))
for (i in str:end){
tell(x, out[,i])
print(x)  
plot(x,
     main = paste("Time = ",4 + (i - str)*0.5, "hr"))  
}
```

Convergence check

```{r, eval=T}
# Create data frame
for (i in 3:11){
  x <- morris(model = NULL, factors = factors, r = 2^i, 
              design = list(type = "oat", levels = 6, grid.jump = 3), 
              binf = binf, bsup = bsup)
  dat <- cbind(1, x$X)
  write.table(dat, file = "setpts.out", row.names = F, sep = "\t")
  out <- mcsim(model = "EB.model.R",  input = "EB_setpts.in.R", dir = "modeling/EB")
  
  # Specific time point
  y <- out$Cvtot_1.5
  tell(x, y)
  
  if (i == 3){ X <- apply(abs(x$ee), 2, mean) } else X <- rbind(X, apply(abs(x$ee), 2, mean))
}
X

# Plot
ylim <- range(X)
plot(X[,1], ylim = ylim, type = "b", xaxt = "n")
axis(1, at = seq(3,11,2), labels = 2^seq(5,13,2))
for(i in 2 :11){
  lines(X[,i], type = "b")
}
```


## 5.3 Fourier amplitude sensitivity testing

**Use pksensi and conduct FAST method to find the non-influential parameter for blood concentrations.**

```{R, eval=T}
params <- c("BW", "Pb", "Pl", "Pf", "Pm", "Pvrg", "Ppu", "Pbr", "VmaxC", "VmaxClu", "VmaxCvr")
q <- "qunif"
q.arg <- list(list(min = 0.0358, max = 0.0516),
              list(min = 21.35, max = 85.4),
              list(min = 0.98, max = 3.92),
              list(min = 18.2, max = 72.8),
              list(min = 0.3045, max = 1.218),
              list(min = 0.98, max = 3.92),
              list(min = 0.98, max = 3.92),
              list(min = 0.98, max = 3.92),
              list(min = 1.5975, max = 25.56),
              list(min = 3.35, max = 53.6),
              list(min = 4.35, max = 69.6)) 
```

2. Generate parameter space

```{r, eval=T}
set.seed(1234)
x <- rfast99(params, n = 8000, q = q, q.arg = q.arg, replicate = 1) # convergence size
```

```{r, eval=T}
dim(x$a) # c(Evaluation, replication, parameters)
```

3. Set experiment time-points, output variables, and conditions 

```{r, eval=T}
var <- c("Cvtot")
times <- c(4.0, 4.5, 5, 5.5, 6)
condition <- c("Cppm = NDoses (2, 100, 0, 0, 4 )")
```

4. Modeling

**Note:** There is a problem to define `rtol = 1e-11` and `atol = 1e-13`. Update pksensi to the latest version by `devtools::install_github("nanhung/pksensi")`

```{r, eval=T}
fast <- solve_mcsim(x, mName = "EB.model.R",  params = params, time = times,  vars = var, 
                 condition = condition, rtol = 1e-11,atol = 1e-13)
```

5. Visualization and decision

```{r, eval=T}
plot(fast)
```

```{r, eval=T}
pksim(fast)
```

```{r, eval=T}
check(fast, SI.cutoff = 0.05)
```

```{r, eval=T}
heat_check(fast, SI.cutoff = c(0.05, 0.09, 0.15))
```


## 5.4 Morris vs. FAST

**Compare the sensitivity measures (first order, interaction, total order) for Morris (exercise 2) and FAST (exercise 4).**

```{r, eval=T}
set.seed(1234)
morrx <- morris(model = NULL, factors = factors, r = 512,
            design = list(type = "oat", levels = 5, grid.jump = 3),
            binf = binf, bsup = bsup)
X <- cbind(1, morrx$X)
write.table(X, file = "setpts.out", row.names = F, sep = "\t")
out <- mcsim(model = "EB.model.R",  input = "EB_setpts.in.R", dir = "modeling/EB")

y <- "Cvtot_1.2"
str <- which(names(out) == y)
tell(morrx, out[,str])
mu_star <- apply(abs(morrx$ee), 2, mean)
mu_star
```


```{r, eval=T}
fast$mSI
dim(fast$mSI)
```

```{r, eval=T}
plot(mu_star, fast$mSI[2,,1], ylim=c(0,1))
abline(0.05, 0)
```


## 5.5 Acetaminophen-PBPK model

**Reproduce the published result of acetaminophen-PBPK model by following the vignette in pksensi**. 

https://nanhung.rbind.io/pksensi/articles/pbpk_apap.html

0. Create & compile model file

```{R, eval=T}
pksensi::pbpk_apap_model()
makemcsim(model = "pbpk_apap.model")
```

1. Define arguments
```{r, eval=T}
Tg <- log(0.23)
Tp <- log(0.033)
CYP_Km <- log(130)
SULT_Km_apap <- log(300)
SULT_Ki <- log(526)
SULT_Km_paps <- log(0.5)
UGT_Km <- log(6.0e3)
UGT_Ki <- log(5.8e4)
UGT_Km_GA <-log(0.5)
Km_AG <- log(1.99e4)
Km_AS <- log(2.29e4)

rng <- 1.96 

# Parameter distribution
params <- c("lnTg", "lnTp", "lnCYP_Km","lnCYP_VmaxC",
           "lnSULT_Km_apap","lnSULT_Ki","lnSULT_Km_paps","lnSULT_VmaxC",
           "lnUGT_Km","lnUGT_Ki","lnUGT_Km_GA","lnUGT_VmaxC",
           "lnKm_AG","lnVmax_AG","lnKm_AS","lnVmax_AS",
           "lnkGA_syn","lnkPAPS_syn", "lnCLC_APAP","lnCLC_AG","lnCLC_AS")
q <- "qunif"
q.arg <-list(list(Tg-rng, Tg+rng),
             list(Tp-rng, Tp+rng),
             list(CYP_Km-rng, CYP_Km+rng),
             list(-2., 5.),
             list(SULT_Km_apap-rng, SULT_Km_apap+rng),
             list(SULT_Ki-rng, SULT_Ki+rng),
             list(SULT_Km_paps-rng, SULT_Km_paps+rng),
             list(0, 10),
             list(UGT_Km-rng, UGT_Km+rng),
             list(UGT_Ki-rng, UGT_Ki+rng),
             list(UGT_Km_GA-rng, UGT_Km_GA+rng),
             list(0, 10),
             list(Km_AG-rng, Km_AG+rng),
             list(7., 15),
             list(Km_AS-rng, Km_AS+rng),
             list(7., 15),
             list(0., 13),
             list(0., 13),
             list(-6., 1),
             list(-6., 1),
             list(-6., 1))

# Mutivariate
vars <- c("lnCPL_APAP_mcgL", "lnCPL_AG_mcgL", "lnCPL_AS_mcgL")
times <- seq(from = 0.1, to = 12.1, by = 0.2)

# Conditions
conditions <- c("mgkg_flag = 1",
                "OralExp_APAP = NDoses(2, 1, 0, 0, 0.001)",
                "OralDose_APAP_mgkg = 20.0")
```

2. Generate Parameter space

```{r, eval=T}
set.seed(1234)
x <- rfast99(params = params, n = 1000, q = q, q.arg = q.arg, replicate = 1) 
```

3. Modeling

```{r, eval=T}
system.time(out <- solve_mcsim(x, mName = "pbpk_apap.model", 
                               params = params,
                               vars = vars,
                               time = times,
                               condition = conditions))
```


4. Visualization & Decision

```{r eval=T, fig.height=13, fig.width=13}
plot(out, vars = "lnCPL_AG_mcgL")
```

```{r eval=T, fig.height=9, fig.width=9}
heat_check(out)
```

```{r eval=T}
check(out)
```

```{r, include=FALSE}
clear()
```
